Debugging by {\huge\bf\textit{lastChange}}

### 2011-03-09 trunk/papers/fse2011/sig-alternate.tex

By browsing through the code or other means\cite{Barton},
 the developer determines that the value displayed on the 
button is set at line 22. Since the displayed value is incorrect we know 
the bug occured before we hit this line.
To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, execution is paused at line
22. Figure~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is zero. We expected this value to be \texttt{2}.

@@

Looking at line 29, it seems that something is wrong with
\texttt{myCondition.value} which causes line 29 execution. The
developer examines \texttt{myCondition.value} and it is
\texttt{undefined}. The next step is to know when this property got
this value. To do so, the developer runs the \textit{lastChange} command
on \texttt{myCondition.value} at this point. The debugger re-executes the
program and breaks again on line 22, analyzes its queries and shows the developer line 25-the
place \texttt{oldValue} is assigned to
\texttt{myCondition.value}. (If the developer asks for \textit{lastChange} on \texttt{oldValue}, 
the debugger can notify the developer that this variable is never assigned a value).
 Now it is clear that the bug occurs because \texttt{oldValue} is
\texttt{undefined} once execution reaches line 25 (Figure~\ref{fig:example4}).

@@

To see how these functions work, suppose the developer asks for the
last change of \texttt{bar.foo} at the reproduction point in a
program. The debugger calls \texttt{setPropertyChangeHook()} with
\texttt{foo} as the property name and re-executes the
program. Whenever \texttt{foo} changes and the callback function is to
be called, debugger first calls \texttt{objectId()} on the
\texttt{foo} owner object. Then it stores this owner id, the stack
frame locations, and other state values in scope at the call point. 
Then the callback returns to continue execution. Thus the query is not 
a breakpoint in the sense of pausing for user interaction, but breakpoint 
technology can be used to implement the query.
Whenever the execution reaches the reproduction point the debugger
looks at the history of \texttt{foo} changes and finds the last
\texttt{foo} change with the same object id as \texttt{bar} id at the
reproduction point. Figure~\ref{fig:foo-changes1} shows the list of
property \texttt{foo} change events in a hypothetical
execution. \texttt{bar} id at the reproduction point is 1010, so the
last change of \texttt{bar.foo} is the fourth column.

@@

\subsection{{\normalsize\bf\texttt{objectId()}} operation}
\texttt{objectId(obj)} first checks the argument \texttt{obj} for a property \texttt{\_objectId}.
It returns the value if this property is already defined,
otherwise it generates a new id and sets this property. The value of the id is simply an integer incremented 
for each new \texttt{\_objectId} needed. This use of a regular object property  might change
the program behaviour. For example it will apear as an extra property in 
\texttt{for(property in object)} loops. 
The next version of Firefox\footnote[3]{https://developer.mozilla.org/en/JavaScript/Reference/Global\_Objects\\/Object/defineProperty} provides an alternative via \texttt{defineProperty} a standard JavaScript
function which receives 
an argument specifying whether the property is enumerable or not. By setting the enumerable
field \texttt{false} for \texttt{\_objectId}, this property will not
apear in \texttt{for(property in object)} loops and therefore has no
effect on the program execution. Note that \texttt{\_objectId} is
not set for all objects but only those objects that need an id.

>>>>

By browsing through the code or other means\cite{Barton},
 the developer determines that the value displayed on the 
button is set at line 22. Since the displayed value is incorrect we know 
the bug occured before we hit this line.
To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, the execution is paused at line
22. Figure~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is zero. We expected this value to be \texttt{2}.

@@

Looking at line 29, it seems that something is wrong with
\texttt{myCondition.value} which causes line 29 execution. The
developer examines \texttt{myCondition.value} and it is
\texttt{undefined}. The next step is to know when this property got
this value. To do so, the developer runs the \textit{lastChange} command
on \texttt{myCondition.value} at this point. The debugger re-executes the
program and breaks again on line 22, analyzes its queries and shows the developer line 25-the
place \texttt{oldValue} is assigned to
\texttt{myCondition.value}. (If the developer asks for \textit{lastChange} on \texttt{oldValue}, 
the debugger can notify the developer that this variable is never assigned a value).
 Now it is clear that the bug occurs because \texttt{oldValue} is
\texttt{undefined} once the execution reaches line 25 (Figure~\ref{fig:example4}).

@@

To see how these functions work, suppose the developer asks for the
last change of \texttt{bar.foo} at the reproduction point in a
program. The debugger calls \texttt{setPropertyChangeHook()} with
\texttt{foo} as the property name and re-executes the
program. Whenever \texttt{foo} changes and the callback function is to
be called, debugger first calls \texttt{objectId()} on the
\texttt{foo} owner object. Then it stores this owner id, the stack
frame locations, and other state values in scope at the call point. 
Then the callback returns to continue the execution. Thus the query is not 
a breakpoint in the sense of pausing for user interaction, but breakpoint 
technology can be used to implement the query.
Whenever the execution reaches the reproduction point the debugger
looks at the history of \texttt{foo} changes and finds the last
\texttt{foo} change with the same object id as \texttt{bar} id at the
reproduction point. Figure~\ref{fig:foo-changes1} shows the list of
property \texttt{foo} change events in a hypothetical
execution. \texttt{bar} id at the reproduction point is 1010, so the
last change of \texttt{bar.foo} is the fourth column.

@@

\subsection{{\normalsize\bf\texttt{objectId()}} operation}
\texttt{objectId(obj)} first checks the argument \texttt{obj} for a property \texttt{\_objectId}.
It returns the value if this property is already defined,
otherwise it generates a new id and sets this property. The value of the id is simply an integer incremented 
for each new \texttt{\_objectId} needed. This use of a regular object property  might change
the program behaviour. For example it will apear as an extra property in 
\texttt{for(property in object)} loops. 
The next version of Firefox\footnote[3]{https://developer.mozilla.org/en/JavaScript/Reference\\/Global\_Objects/Object/defineProperty} provides an alternative via \texttt{defineProperty} a standard JavaScript
function which receives 
an argument specifying whether the property is enumerable or not. By setting the enumerable
field \texttt{false} for \texttt{\_objectId}, this property will not
apear in \texttt{for(property in object)} loops and therefore has no
effect on the program execution. Note that \texttt{\_objectId} is
not set for all objects but only those objects that need an id.
